---
title: "Exercise 1 - Full Document"
author: "Bertram Melix"
date: "2023-09-23"
output: html_document
---

```{r}
library(dplyr)
library(knitr)
library(gt)
library(tidyverse)
library(haven)
library(skimr)
library(naniar)
library(Hmisc)
library(sjlabelled)
library(gt)
library(gtsummary)
library(table1)
```


# General Social Survey - Family (GSS) - Dataset includes 20,602 observations and 1,316 variables. 
```{r}
gss <- read.csv("C:/Users/bertr/Desktop/CAnD3 Program Application/Exercise1/CAND3 Data/CAND3 Data/GSS 2017 - Family, Cycle 31/gss-12M0025-E-2017-c-31_F1.csv")
```


# Here we filter to keep relevant data points for the analysis - omitting obervations where respondents did not answer, responded with "dont know" or "Valid Skip" unless otherwise noted. 
```{r}
gss_clean <- gss %>%
  filter(
    SRH_110 <= 5, # Self Rated Health - Excellent (1) to Poor (5)
    TTLINCG2 <= 6, # Income of Respondent - <$25,000 (1) to $125,000 or more (5) 
    HSDSIZEC <= 6, # Household Size of Respondent - One person household (1) to Six or more person household (6)
    LUC_RST <= 3, # Population center indicator - Larger urban population centres (1), Rural areas and small population centres (2), Prince Edward Island (3)
    GU_110 <= 3, # Birth or adoption - Yes (both birth parents) (1), Yes (both adoptive parents (2), No (3)
    APARDIVC <= 97, # Age of respondent when parents divorced - capped at age 35 and older (0-35yrs), Valid Skip (96), &  Don't Know (97)
    ODR_10 <= 2, # Owned or Renter - Owned by (1) and Rented (2)
    UHW_16GR <= 6, # Average number of hours worked/week - 0 hour (1) to 50.1 hours and more (5), Valid Skip (6)? 
    BRTHMACR <= 6 # Place of birth of respondent - Americas (1), Europe (2), Africa (3), Asia (4), Oceania and others (5)
  ) %>%
  select(SRH_110, TTLINCG2, HSDSIZEC, LUC_RST, GU_110, UHW_16GR, APARDIVC, ODR_10, BRTHMACR)  %>%
rename(
    srh = SRH_110, 
    totinc = TTLINCG2, 
    hhsize = HSDSIZEC, 
    rururb = LUC_RST, 
    adopt = GU_110, 
    avghw = UHW_16GR, 
    divorce = APARDIVC, 
    hmown = ODR_10, 
    bplace = BRTHMACR
  )
```



# Create a summary table with mean and standard deviation for each variable
```{r}
summary_table <- gss_clean %>%
  summarise(
    `Self-rated Health` = mean(srh),
    `Income of Respondent` = mean(totinc),
    `Household Size` = mean(hhsize),
    `Rural/Urban` = mean(rururb),
    `Family Living Situation` = mean(adopt),
    `Average Number of Hours Worked/Week` = mean(avghw),
    `Dwelling - Owned or Rented` = mean(hmown),
    `Place of Birth of Respondent` = mean(bplace)
  ) %>%
  gather(key = "Variable", value = "Mean") %>%
  arrange(Variable)

# calculate mean and std dev for divorce variable (exclude values greater than 35)
divorce_summary <- gss_clean %>%            
  filter(divorce <= 35) %>%
  summarise(
    `Age of Respondent When Parents Divorced` = mean(divorce, na.rm = TRUE)
  ) %>%
  gather(key = "Variable", value = "Mean")


summary_table <- bind_rows(summary_table, divorce_summary)


std_dev_table <- gss_clean %>%
  summarise(
    `Self-rated Health` = sd(srh),
    `Income of Respondent` = sd(totinc),
    `Household Size` = sd(hhsize),
    `Rural/Urban` = sd(rururb),
    `Family Living Situation` = sd(adopt),
    `Age of Respondent When Parents Divorced` = sd(divorce),
    `Average Number of Hours Worked/Week` = sd(avghw),
    `Dwelling - Owned or Rented` = sd(hmown),
    `Place of Birth of Respondent` = sd(bplace)
  ) %>%
  gather(key = "Variable", value = "Std. Dev.") %>%
  arrange(Variable)

the_table <- merge(summary_table, std_dev_table, by = "Variable")


kable(the_table, caption = "Table 1. Descriptive Statistics for select variables from the General Social Survey, Cycle 31, 2017: Family")
```



# Binary recoding
```{r}
gss_clean <- gss_clean %>%
  mutate(
    adopt = ifelse(adopt == 3, 1, 0), # Setting values 'No' as 1, and 'Yes' (adoptive or birth) to 0
    hmown = ifelse(hmown == 2, 1, 0), # Setting values 'Rented' as 1, 'Owned by' as 0
  )
```


# Nominal Recoding
```{r}
gss_clean <- gss_clean %>%
  mutate(bplace = as.factor(bplace),
    rururb = as.factor(rururb)
  )
```


# Formula for OLS
```{r}
srh.form <- srh ~ hhsize + rururb + adopt + divorce + avghw + hmown + bplace + totinc
```


# OLS
```{r}
srh.lm<- lm(srh.form, data= gss_clean)
summary(srh.lm)
```

# Table 2 - OLS results (Self Rated Health)
```{r}
srh_results <- summary(srh.lm)$coefficients 

# Convert matrix to data frame and add Variable column from row names
regression_table <- as.data.frame(srh_results)
regression_table$Variable <- rownames(srh_results)

# Reorder the columns
regression_table <- regression_table %>%
  select(Variable, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`)

# View the table
regression_table

# Display the table using kable
knitr::kable(regression_table, caption = "Table 2. Linear Regression Results for SRH on select variables from the General Social Survey, Cycle 31, 2017: Family")

```


-----------------------------------------------------------------------------------------------------------
# Code Graveyard


# Ordinal Recoding for Ordinal Logistic Regression
```{r}
library(dplyr)

gss_clean <- gss_clean %>%
  mutate(
    srh = ordered(srh, levels = c(1, 2, 3, 4, 5)),
    totinc = ordered(totinc, levels = c(1, 2, 3, 4, 5, 6)),
    avghw = ordered(avghw, levels = c(1, 2, 3, 4, 5, 6))
  )
```



```{r}
library(MASS)

srh.olr <- polr(srh.form, data = gss_clean, Hess = TRUE)
summary(srh.olr)
```

